---
title: "Assignment 7"
author: "Nathan Underwood"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r Initialization}
library(rvest)
library(tidyverse)
```

## Question 1

At the Insurance Institute for Highway Safety, they have [data](https://www.iihs.org/topics/fatality-statistics/detail/state-by-state) about human fatalities in vehicle crashes. From this web page, import the data from the Fatal Crash Totals data table and produce a bar graph gives the number of deaths per 100,000 individuals. Be sure to sort the states by highest to lowest mortality.
_Hint: If you have a problem with the graph being too squished vertically, you can set the chunk options `fig.height` or `fig.width` to make the graph larger, but keeping the font sizes the same. The result is that the text is more spread apart. The chunk options `out.height` and `out.width` shrink or expand everything in the plot. By making the `fix.XXX` options large and `out.XXX` options small, you are effectively decreasing the font size of all the elements in the graph. The other trick is to reset the font size using a theme `element_text` option: `theme(text = element_text(size = 9))`._

```{r Question 1, fig.width = 20, fig.wdith = 20}
url <- "https://www.iihs.org/topics/fatality-statistics/detail/state-by-state"
page <- read_html(url)

fatalCrashTotal <- page %>% 
  html_nodes("table") %>% 
  .[[1]] %>% # grab the first available table
  html_table(header = FALSE, fill = TRUE) %>% # convert the HTML table into a data frame
  slice(-1 * 1:2 )
  
# rename the data
fatalCrashTotal <- fatalCrashTotal %>%
  magrittr::set_colnames(c("State", "Population", "Vehicle Miles Traveled (Milions)", "Fatal Crashes", "Deaths", "Deaths per 100,000 Population", "Deaths per 100 Million Vehicles Miles Traveled"))

# clean the data
fatalCrashTotal <- fatalCrashTotal %>% 
  mutate(across(2:7, str_remove_all, ',')) %>% 
  mutate(across(2:7, as.numeric)) %>% # correct the data types
  filter(!(State %in% "U.S. total")) %>% 
  mutate(State = fct_reorder(`State`, `Deaths per 100,000 Population`))
  
head(fatalCrashTotal)

# plot the data in a bar graph
ggplot(fatalCrashTotal, aes(x = `State`, y = `Deaths per 100,000 Population`)) +
  geom_col() +
  coord_flip() +
  labs(title='Fatality Rate per 100 Thousand People by State')
```

## Question 2

From the same IIHS website, import the data about seat belt use. Join the Fatality data with the seat belt use and make a scatter plot of percent seat belt use vs number of fatalities per 100,000 people.

```{r Question 2}
seatBeltUse <- page %>% 
  html_nodes("table") %>% 
  .[[5]] %>% # grab the fourth table
  html_table(header = FALSE, fill = TRUE) %>% 
  slice(-1 * 1:2 )

# rename the data
seatBeltUse <- seatBeltUse %>% 
  magrittr::set_colnames(c("State", "Population", "Vehicle Miles Traveled (Milions)", "Fatal Crashes", "Deaths", "Deaths per 100,000 Population", "Deaths per 100 Million Vehicles Miles Traveled"))

# clean the data
seatBeltUse <- seatBeltUse %>% 
  mutate(across(2:6, str_remove_all, ',')) %>% 
  # mutate(across(2:6, as.numeric)) %>% # correct the data types
  filter(!(State %in% "U.S. total"))
  # mutate(State = fct_reorder(`State`, `Deaths per 100,000 Population`))

# merge the fatalities with seat belt use
seatBeltFatalities <- fatalCrashTotal %>% 
  cbind(seatBeltUse)
```

