---
title: "Assignment4"
author: "Nathan Underwood"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r Initialization, include=FALSE}
library(tidyverse)
library(lubridate)
```

## Question 1

 A common task is to take a set of data that has multiple categorical 
    variables and create a table of the number of cases for each combination.
    An introductory statistics textbook contains a dataset summarizing student
    surveys from several sections of an intro class. The two variables of
    interest for us are `Gender` and `Year` which are the students gender and
    year in college.
    a)  Download the dataset and correctly order the `Year` variable using the
        following:
        
```{r Question 1-A}
Survey <- read.csv('https://www.lock5stat.com/datasets3e/StudentSurvey.csv', na.strings=c('',' '))
```
    b)  Using some combination of `dplyr` functions, produce a data set with
        eight rows that contains the number of responses for each gender:year
        combination. Make sure your table orders the `Year` variable in the
        correct order of `First Year`, `Sophmore`, `Junior`, and then `Senior`.
        *You might want to look at the following functions: `dplyr::count` and*
        *`dplyr::drop_na`.* 
        
```{r Question 1-B}
SurveyClean <- Survey %>% 
  mutate(Year = fct_relevel(Year, "FirstYear", "Sophomore", "Junior", "Senior")) %>% 
  drop_na() %>% 
  count(Sex, Year)
```
    c)  Using `tidyr` commands, produce a table of the number of responses in
        the following form:
        
        |   Gender    |  First Year  |  Sophmore  |  Junior   |  Senior   |
        |:-----------:|:------------:|:----------:|:---------:|:---------:|
        |  **Female** |              |            |           |           |  
        |  **Male**   |              |            |           |           | 
        
```{r Question 1-C}
SurveyTable <- SurveyClean %>% 
  pivot_wider(names_from = Year, values_from = n)

SurveyTable
```

## Question 2

From the book website, there is a .csv file of the daily maximum temperature
    in Flagstaff at the Pulliam Airport. The direction link is at:
    [https://raw.githubusercontent.com/dereksonderegger/444/master/data-raw/FlagMaxTemp.csv](https://raw.githubusercontent.com/dereksonderegger/444/master/data-raw/FlagMaxTemp.csv)
    a)  Create a line graph that gives the daily maximum temperature for 2005.
        *Make sure the x-axis is a date and covers the whole year.*
        
```{r Question 2-A}
DailyMaxTemps <- read.csv('https://raw.githubusercontent.com/dereksonderegger/444/master/data-raw/FlagMaxTemp.csv', na.strings=c('',' '))

MaxTemps2005 <- DailyMaxTemps %>% 
  filter(Year == "2005") %>%
  pivot_longer(names_to = "Day", values_to = "Temps", X1:X31) %>%
  mutate(Day = str_replace(Day, pattern = 'X', replacement = '')) %>%
  mutate(Date = make_date(year = Year, month = Month, day = Day)) %>% 
  mutate(Temps = as.numeric(Temps)) %>% 
  drop_na
  
  
ggplot(MaxTemps2005, aes(x = Date, y = Temps)) +
  geom_point() +
  geom_line()
```
    b)  Create a line graph that gives the monthly average maximum temperature
        for 2013 - 2015. *Again the x-axis should be the date and the axis*
        *spans 3 years.*
        
```{r Question 2-B}
MaxTemps2013_15 <- DailyMaxTemps %>% 
  filter(Year >= 2013, Year <= 2015) %>% 
  pivot_longer(names_to = "Day", values_to = "Temps", X1:X31) %>%
  mutate(Day = str_replace(Day, pattern = 'X', replacement = '')) %>%
  mutate(Temps = as.numeric(Temps)) %>% 
  group_by(Year, Month) %>% 
  drop_na %>% 
  summarise(mean = mean(Temps)) %>% 
  mutate(Date = make_date(year = Year, month = Month))

ggplot(MaxTemps2013_15, aes(x = Date, y = mean)) +
  geom_point() +
  geom_line()
```
## Question 4

For this problem we will consider two simple data sets.

```{r, Question 4-Data}
A <- tribble(
  ~Name, ~Car,
  'Alice', 'Ford F150',
  'Bob',   'Tesla Model III',
  'Charlie', 'VW Bug')

B <- tribble(
  ~First.Name, ~Pet,
  'Bob',  'Cat',
  'Charlie', 'Dog',
  'Alice', 'Rabbit')
```
    a)  Squish the data frames together to generate a data set with three rows
        and three columns. Do two ways: first using `cbind` and then using one
        of the `dplyr` `join` commands.
        
```{r Question 4-A}
B <- rename(B, Name = First.Name)

C <- cbind(A, B)

C <- full_join(A, B)

C
```
    b)  It turns out that Alice also has a pet guinea pig. Add another row to
        the `B` data set. Do this using either the base function `rbind`, or
        either of the `dplyr` functions `add_row` or `bind_rows`.
        
```{r Question 4-B}
B <- add_row(B, Name = "Alice", Pet = "Guinea Pig")

B
```
    c)  Squish the `A` and `B` data sets together to generate a data set with
        four rows and three columns. Do this two ways: first using `cbind` and
        then using one of the `dplyr` `join` commands. Which was easier to
        program? Which is more likely to have an error.
        
```{r Question 4-C}
# D <- cbind(A, B)

D <- full_join(A, B)

D
```
The `cbind` implementation is more likely to have errors because it struggles when 
there are tables of different sizes, unlike join.

## Question 5

Data table joins are extremely common because effective database design
    almost always involves having multiple tables for different types of objects.
    To illustrate both the table joins and the usefulness of multiple tables we
    will develop a set of data frames that will represent a credit card company's
    customer data base. We will have tables for Customers, Retailers, Cards, and
    Transactions.  Below is code that will create and populate these tables.
    
```{r Question 5-Data}
Customers <- tribble(
  ~PersonID, ~Name, ~Street, ~City, ~State,
  1, 'Derek Sonderegger',  '231 River Run', 'Flagstaff', 'AZ',
  2, 'Aubrey Sonderegger', '231 River Run', 'Flagstaff', 'AZ',
  3, 'Robert Buscaglia', '754 Forest Heights', 'Flagstaff', 'AZ',
  4, 'Roy St Laurent', '845 Elk View', 'Flagstaff', 'AZ')

Retailers <- tribble(
  ~RetailID, ~Name, ~Street, ~City, ~State,
  1, 'Kickstand Kafe', '719 N Humphreys St', 'Flagstaff', 'AZ',
  2, 'MartAnnes', '112 E Route 66', 'Flagstaff', 'AZ',
  3, 'REI', '323 S Windsor Ln', 'Flagstaff', 'AZ' )

Cards <- tribble(
  ~CardID, ~PersonID, ~Issue_DateTime, ~Exp_DateTime,
  '9876768717278723',  1,  '2019-9-20 0:00:00', '2022-9-20 0:00:00',
  '5628927579821287',  2,  '2019-9-20 0:00:00', '2022-9-20 0:00:00',
  '7295825498122734',  3,  '2019-9-28 0:00:00', '2022-9-28 0:00:00',
  '8723768965231926',  4,  '2019-9-30 0:00:00', '2022-9-30 0:00:00' ) 

Transactions <- tribble(
  ~CardID, ~RetailID, ~DateTime, ~Amount,
  '9876768717278723', 1, '2019-10-1 8:31:23',    5.68,
  '7295825498122734', 2, '2019-10-1 12:45:45',  25.67,
  '9876768717278723', 1, '2019-10-2 8:26:31',    5.68,
  '9876768717278723', 1, '2019-10-2 8:30:09',    9.23,
  '5628927579821287', 3, '2019-10-5 18:58:57',  68.54,
  '7295825498122734', 2, '2019-10-5 12:39:26',  31.84,
  '8723768965231926', 2, '2019-10-10 19:02:20', 42.83) 

Cards <- Cards %>% 
  mutate( Issue_DateTime = lubridate::ymd_hms(Issue_DateTime),
          Exp_DateTime   = lubridate::ymd_hms(Exp_DateTime) )

Transactions <- Transactions %>% 
  mutate( DateTime = lubridate::ymd_hms(DateTime))
```
    a)  Create a table that gives the credit card statement for Derek. It should
        give all the transactions, the amounts, and the store name. Write your
        code as if the only initial information you have is the customer's name.
        *Hint: Do a bunch of table joins, and then filter for the desired customer*
        *name. To be efficient, do the filtering first and then do the table joins.*
        
```{r Question 5-A}
DerekStatement <- Customers %>% 
  filter(Name == "Derek Sonderegger") %>% 
  inner_join(Cards) %>% 
  inner_join(Transactions) %>% 
  inner_join(Retailers, by = "RetailID")

DerekStatement
```
    b)  Aubrey has lost her credit card on Oct 15, 2019. Close her credit card at
        4:28:21 PM and issue her a new credit card in the `Cards` table.
        *Hint: Using Aubrey's name, get necessary CardID and PersonID and save*
        *those as `cardID` and `personID`. Then update the `Cards` table row that*
        *corresponds to the `cardID` so that the expiration date is set to the time*
        *that the card is closed. Then insert a new row with the `personID` for*
        *Aubrey and a new `CardID` number that you make up.* 
        
```{r Question 5-B}
findPersonInfo <- function(name) {
  if(name %in% Customers$Name) 
  {
      dataRow <- which(Customers$Name == name, arr.ind = TRUE)
      personID <- Customers$PersonID[dataRow]
  }
}

findCardInfo <- function(name) {
  if(name %in% Customers$Name) 
  {
    dataRow <- which(Cards$PersonID == personID, arr.ind = TRUE)
    cardID <- Cards$CardID[dataRow]
  }
}

personID <- findPersonInfo("Aubrey Sonderegger")
cardID <- findCardInfo("Aubrey Sonderegger")

Cards$Exp_DateTime[2] <- ymd_hms("2019-9-15 4:28:21 PM")

Cards <- Cards %>% 
  add_row(CardID = "7742900840178223", PersonID = personID, 
          Issue_DateTime = ymd("2019-9-15"), Exp_DateTime = ymd("2022-9-15"))

Cards
```
The `findCardInfo` function needs an additional check to see if the card it is
accessing is a valid card, rather than an expired one.   
    c)  Aubrey is using her new card at Kickstand Kafe on Oct 16, 2019 at
        2:30:21 PM for coffee with a charge of $4.98. Generate a new transaction
        for this action. 
        *Hint: create temporary variables `card`,`retailid`,`datetime`, and*
        *`amount` that contain the information for this transaction and then*
        *write your code to use those. This way in the next question you can just*
        *use the same code but modify the temporary variables. Alternatively, you*
        *could write a function that takes in these four values and manipulates the*
        *tables in the GLOBAL environment using the `<<-` command to assign a result*
        *to a variable defined in the global environment. The reason this is OK is*
        *that in a real situation, these data would be stored in a database and we*
        *would expect the function to update that database.*
        
```{r Question 5-C}
card <- "7742900840178223"
retailID <- 1
dateTime <- ymd_hms("2019-10-16 2:30:21 PM")
amount <- 4.98

ValidCards <- Cards %>% 
  filter(CardID == card, Issue_DateTime <= dateTime, dateTime <= Exp_DateTime)

if( card %in% ValidCards$CardID) {
  
  Transactions <- Transactions %>% 
    add_row(CardID = card, RetailID = retailID, DateTime = dateTime, Amount = amount)
  
} else {
  
  print('Card Denied')
  
}

Transactions
```
    d)  On Oct 17, 2019, some nefarious person is trying to use her OLD credit
        card at REI. Make sure your code in part (c) first checks to see if the
        credit card is active before creating a new transaction. Using the same
        code, verify that the nefarious transaction at REI is denied.
        *Hint: your check ought to look something like this:*

```{r Question 5-D}
card <- "5628927579821287"
retailID <- 3
dateTime <- ymd_hms("2019-10-17 3:23:19 PM")
amount <- 374.63

if( card %in% ValidCards$CardID) {
  
  Transactions <- Transactions %>% 
    add_row(CardID = card, RetailID = retailID, DateTime = dateTime, Amount = amount)
  
} else {
  
  print('Card Denied')
  
}
```
    e)  Generate a table that gives the credit card statement for Aubrey. It
        should give all the transactions, amounts, and retailer name for both
        credit cards she had during this period.

```{r Question 5-E}
AubreyStatement <- Customers %>% 
  filter(Name == "Aubrey Sonderegger") %>% 
  inner_join(Cards) %>% 
  inner_join(Transactions) %>% 
  inner_join(Retailers, by = "RetailID")

AubreyStatement
```



