---
title: "Assignment 3"
author: "Nathan Underwood"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r Initialization, include=FALSE}
library(tidyverse)
```

# Chapter 11

## Question 1

For the following regular expression, explain in words what it matches on. Then add test strings to demonstrate that it in fact does match on the pattern you claim it does. Make sure that your test set of strings has several examples that match as well as several that do not. *If you copy the Rmarkdown code for these exercises directly from my source pages, make sure to remove the `eval=FALSE` from the R-chunk headers.*
    a) This regular expression matches any string that contains the letter 'a' in it.
    
```{r Question 1-A (Chapter 11)}
strings <- c("123", "abc", "this is a string", "a_b_c")
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, 'a') )
```
    b) This regular expression matches any string that contains the letters 'a' and 'b' right next to each other.
    
```{r Question 1-B (Chapter 11)}
strings <- c("123", "abc", "this is a string", "a_b_c")
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, 'ab') )
```
    c)  This regular expression matches any string that contains either an 'a', 'b', or 'c'.
    
```{r Question 1-C (Chapter 11)}
strings <- c("123", "abc", "this is a string", "a_b_c")
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, '[ab]') )
```
    d)  This regular expression matches any string that starts with an 'a', 'b', or 'c'.
    
```{r Question 1-D (Chapter 11)}
strings <- c("123", "abc", "this is a string", "a_b_c")
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, '^[ab]') )
```
    e)  This regular expression matches any string that has the following in this particular order: 1) at least one number; 2) exactly one white space; and 3) and an 'a' or an 'A'.
    
```{r Question 1-E (Chapter 11)}
strings <- c("11 a", "   aA1", "   1aA1", "1 aA", "1 1aA", "11 aA")
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, '\\d+\\s[aA]') )
```
    f)  This regular expression matches any string that has the following in this particular order: 1) at least one number; and 2) and an 'a' or an 'A'. Optionally, there can be white spaces between the digits and the letters.
    
```{r Question 1-F (Chapter 11)}
strings <- c("11a", "   aA 1", "   1aA1", "1 aA", "1 1aA", "1 1")
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, '\\d+\\s*[aA]') )
```
    g)  This regular expression matches any string with any number of characters (including zero!).
    
```{r Question 1-G (Chapter 11)}
strings <- c('', "abcdefg", "1000", "aljsfhdg.2uig4")
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, '.*') )
```
    h) This regular expression matches any string that begins with two alphanumerical characters followed by the word "bar". This string must be five characters in length. 
    
```{r Question 1-H (Chapter 11)}
strings <- c("1Abar", "1bar", "bar", "babar")
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, '^\\w{2}bar') )
```
    i) This regular expression matches any string that is either 1) a string from part H; or 2) or the word "foobar" with a period stuck in the middle.
    
```{r Question 1-I (Chapter 11)}
strings <- c("foo.bar", "foolbar", "fobar", "1bar", "bar", "babar")
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, '(foo\\.bar)|(^\\w{2}bar)') )
```

## Question 2
The following file names were used in a camera trap study. The S number represents the site, P is the plot within a site, C is the camera number within the plot, the first string of numbers is the YearMonthDay and the second string of numbers is the HourMinuteSecond.

```{r Question 2 Date Frame (Chapter 11)}
file.names <- c( 'S123.P2.C10_20120621_213422.jpg',
                 'S10.P1.C1_20120622_050148.jpg',
                 'S187.P2.C2_20120702_023501.jpg')
```
    Produce a data frame with columns corresponding to the `site`, `plot`, `camera`, `year`, `month`, `day`, `hour`, `minute`, and `second` for these three file names. So we want to produce code that will create the data frame:
    
```{r, Question 2-A (Chapter 11)}
file <- 
  str_replace_all(file.names, pattern = '_', replacement = '.') %>% 
  str_split_fixed(pattern = '\\.', n = 6)

Year <- str_sub(file[,4], start = 1, end = 4)
Month <- str_sub(file[,4], start = 5, end = 6)
Day <- str_sub(file[,4], start = 7, end = 8)
Hour <- str_sub(file[,5], start = 1, end = 2)
Minute <- str_sub(file[,5], start = 3, end = 4)
Second <- str_sub(file[,5], start = 5, end = 6)

file <- data.frame(file) %>% 
  select(-X4) %>% 
  select(-X5) %>% 
  select(-X6) %>% 
  mutate(Year = Year) %>% 
  mutate(Month = Month) %>% 
  mutate(Day = Day) %>% 
  mutate(Hour = Hour) %>% 
  mutate(Minute = Minute) %>% 
  mutate(Second = Second) %>% 
  rename(Site = X1) %>% 
  rename(Plot = X2) %>% 
  rename(Camera = X3)
  
head(file)
```
 
## Question 3
The full text from Lincoln's Gettysburg Address is given below. Calculate the mean word length *Note: consider 'battle-field' as one word with 11 letters*).

```{r Question 3}
Gettysburg <- 'Four score and seven years ago our fathers brought forth on this 
continent, a new nation, conceived in Liberty, and dedicated to the proposition 
that all men are created equal.

Now we are engaged in a great civil war, testing whether that nation, or any 
nation so conceived and so dedicated, can long endure. We are met on a great 
battle-field of that war. We have come to dedicate a portion of that field, as 
a final resting place for those who here gave their lives that that nation might 
live. It is altogether fitting and proper that we should do this.

But, in a larger sense, we can not dedicate -- we can not consecrate -- we can 
not hallow -- this ground. The brave men, living and dead, who struggled here, 
have consecrated it, far above our poor power to add or detract. The world will 
little note, nor long remember what we say here, but it can never forget what 
they did here. It is for us the living, rather, to be dedicated here to the 
unfinished work which they who fought here have thus far so nobly advanced. It 
is rather for us to be here dedicated to the great task remaining before us -- 
that from these honored dead we take increased devotion to that cause for which 
they gave the last full measure of devotion -- that we here highly resolve that 
these dead shall not have died in vain -- that this nation, under God, shall 
have a new birth of freedom -- and that government of the people, by the people, 
for the people, shall not perish from the earth.'

GettysburgSplit <-
  Gettysburg %>% 
  str_replace_all(pattern = "\\.|\\,|\\\n|-- " ,replacement = '') %>% 
  str_split(pattern = ' ')

str_length(GettysburgSplit)

meanLength <- (str_length(GettysburgSplit) - 1) / 269 / 2

meanLength
```
# Chapter 12

## Question 1

Convert the following to date or date/time objects.
    a) September 13, 2010.
    
```{r Question 1-A (Chapter 12)}

```
    b) Sept 13, 2010.
    
```{r Question 1-B (Chapter 12)}

```
    c) Sep 13, 2010.
    
```{r Question 1-C (Chapter 12)}

```
    d) S 13, 2010. Comment on the month abbreviation needs.
    
```{r Question 1-D (Chapter 12)}

```
    e) 07-Dec-1941.
    
```{r Question 1-E (Chapter 12)}

```
    f) 1-5-1998. Comment on why you might be wrong.
    
```{r Question 1-F (Chapter 12)}

```
    g) 21-5-1998. Comment on why you know you are correct.
    
```{r Question 1-G (Chapter 12)}

```
    h) 2020-May-5 10:30 am
    
```{r Question 1-H (Chapter 12)}

```
    i) 2020-May-5 10:30 am PDT (ex Seattle)
    
```{r Question 1-I (Chapter 12)}

```
    j) 2020-May-5 10:30 am AST (ex Puerto Rico)
    
```{r Question 1-J (Chapter 12)}

```

## Question 2

Using just your date of birth (ex Sep 7, 1998) and today's date calculate the following _Write your code in a manner that the code will work on any date after you were born._:
    a) Calculate the date of your 64th birthday.
    b) Calculate your current age (in years). _Hint: Check your age is calculated correctly if your birthday was yesterday and if it were tomorrow!_ 
    d) Using your result in part (b), calculate the date of your next birthday.
    e) The number of _days_ until your next birthday.
    f) The number of _months_ and _days_ until your next birthday.